Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2019-02-12 Manuscript/manuscript.Rmd
manuscript edits

In addition to employing suitable metrics for evaluating models validated on imbalanced data, there also exist several algorithmic or data level solutions. Algorithmic level solutions involve modifyng existing classifiers in order to curtail the bias induced by imbalanced classes [@estabrooks2004multiple; @chawla2003smoteboost; @chen2004using; @khoshgoftaar2007empirical]. Data level solutions mostly constitute as resampling techniques, in which random sampling is used to either replicate or removed data points in order to create a more balanced dataset [@chawla2004special; @chawla2009data]. It has been shown that quite simple resampling techniques can drastically improve model performance in the face of imbalanced data [@japkowicz2002class; @dubey2014analysis]. For the purposes of this study, we focus our attention toward data level solutions so that all results may be classifier-independent, and, thus, more easily comparable.

The other aspect of TAD boundary modelling not considered by the current methods is the concept of genomic distance. The typical approaches for modeling the association between TAD boundaries and epigenomic elements include overlap counts [??? REFs], or percent counts [??? Do we have refs for this?]. The use of distances from TAD boundaries to regions of various genomic features, measured in the number of bases between them, has largely been ignored. However, these distance characteristics have the potential to offer a more accurate spatial representation of which genomic features are associated with TAD boundaries.

@@

In this study, we have developed an ensemble framework for predicting TAD boundaries that addresses class imbalance problem. We evaluated Hi-C data for both the GM12878 and K562 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four re-sampling techniques aimed at creating balanced classes. [??? "Following feature reduction using LASSO," - should this be added?] Random forest classification was performed to both compare each combination of the above ensemble, and investigate which functional genomic features were most associated with the TAD boundaries.  Model performance was evaluated using area under the precision-recall curve (AUPRC), F1-Score, and MCC.

>>>>

In addition to employing suitable metrics for evaluating models validated on imbalanced data, there also exist several algorithmic or data level solutions. Algorithmic level solutions involve modifyng existing classifiers in order to curtail the bias induced by imbalanced classes [@estabrooks2004multiple; @chawla2003smoteboost; @chen2004using; @khoshgoftaar2007empirical]. Data level solutions mostly constitute as resampling techniques, in which random sampling is used to either replicate or remove data points in order to create a more balanced dataset [@chawla2004special; @chawla2009data]. It has been shown that quite simple resampling techniques can drastically improve model performance in the face of imbalanced data [@japkowicz2002class; @dubey2014analysis]. For the purposes of this study, we focus our attention toward data level solutions so that all results may be classifier-independent, and, thus, more easily comparable.

The other aspect of TAD boundary modelling not considered by the current methods is the concept of genomic distance. The typical approaches for modeling the association between TAD boundaries and epigenomic elements include overlap counts or percents [??? Do we have refs for this?; !!!should I repeat past used references here?]. Counts refer to the total number of instances where a genomic annotation overlaps with a genomic region of interest, while percents refers to the percent of overlap that occurred (usually in terms of base pairs). The use of distances from TAD boundaries to regions of various genomic features, measured in the number of bases between them, has largely been ignored. However, these distance characteristics have the potential to offer a more accurate spatial representation of which genomic features are associated with TAD boundaries.

@@

In this study, we have developed an ensemble framework for predicting TAD boundaries in order to systematically compare many data driven solutions used to address the issues presented above. We evaluated Hi-C data for both the GM12878 and K562 cell line obtained at 10 kb, 25 kb, 50 kb, and 100 kb resolution. For each resolution, we examined the inclusion of three different predictor types, including distance as the measure of association between genomic elements. We then tested four different re-sampling techniques aimed at creating balanced classes. Noninformative features were removed using a variable reduction algorithm and random forest classifications were performed to assess model performance, and investigate which genomic annotations were most associated with the TAD boundaries for a given system from the complete ensemble.  Model performance was evaluated using area under the precision-recall curve (AUPRC), F1-Score, Youden's index (YI), and Matthew's correlation coefficient (MCC).
